---
title: "Indicators by stock"
author: "Abigail Tyrell"
date: "`r format(Sys.time(), '%d %b %Y')`"
output:
  html_document:
    toc: true
    number_sections: true
    toc_depth: 3
    toc_float:
      collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.height = 7)
`%>%` <- magrittr::`%>%`
```

```{r}
esp_traffic_fig <- function(data,
                             name,
                             out = "one_pager",
                             paginate = FALSE,
                             label = TRUE,
                             caption = "",
                             ncolumn = 1,
                             silent = FALSE,
                             min_year = 1985,
                            sd = FALSE,
                             ...) {
  maxyear <- max(data$YEAR)

  ### data prep ----
  ### write code
  ### colnames being used in current code:
  ### YEAR, DATA_VALUE, mean, sd, name (indicator name), label_num, score,

  dat <- data %>%
    dplyr::group_by(INDICATOR_NAME) %>%
    dplyr::mutate(mean = mean(DATA_VALUE, na.rm = TRUE),
                  sd = sd(DATA_VALUE, na.rm = TRUE),
                  name = stringr::str_wrap(INDICATOR_NAME,
                                           width = 75))

  dat$name <- factor(dat$name, levels = unique(dat$name))

  # base figure ----
  plt <- ggplot2::ggplot(
    dat,
    ggplot2::aes(
      x = .data$YEAR,
      y = .data$DATA_VALUE,
      group = .data$name
    )
  ) +
      # add sd ribbon if specified
  {
    if(sd) {
      ggplot2::geom_ribbon(ggplot2::aes(ymin = .data$DATA_VALUE - .data$DATA_SD,
                                        ymax = .data$DATA_VALUE + DATA_SD),
                           alpha = 0.25,
                           fill = "darkgreen")
    }
  } +
    
    ggplot2::geom_hline(ggplot2::aes(
      yintercept = .data$mean + .data$sd,
      group = .data$name
    ),
    color = "darkgreen",
    linetype = "solid"
    ) +
    ggplot2::geom_hline(ggplot2::aes(
      yintercept = .data$mean - .data$sd,
      group = .data$name
    ),
    color = "darkgreen",
    linetype = "solid"
    ) +
    ggplot2::geom_hline(ggplot2::aes(
      yintercept = .data$mean,
      group = .data$name
    ),
    color = "darkgreen",
    linetype = "dotted"
    ) +
    ggplot2::geom_point() +
    ggplot2::geom_line(data = dat %>%
                         tidyr::drop_na(.data$DATA_VALUE)) +
    ggplot2::ylab("") +
    ggplot2::scale_y_continuous(labels = scales::comma,
                                n.breaks = 4) +
    ggplot2::scale_x_continuous(n.breaks = 4) +
    ggplot2::theme_bw(base_size = 16) +
    ggplot2::theme(strip.text = ggplot2::element_text(size = 10),
                   aspect.ratio = 0.25)
  
  # test linear trend and add line if there is a trend
  pdat <- dat %>%
    tidyr::drop_na(DATA_VALUE) %>%
    dplyr::group_by(name) %>%
    dplyr::mutate(n_obs = dplyr::n()) %>%
    dplyr::filter(n_obs > 30)

  if(nrow(pdat) > 0){
    pdat <- pdat %>%
      # Fit the linear model for each grouping
      dplyr::do(fit = broom::tidy(lm(DATA_VALUE ~ YEAR, data = .))) %>%
      tidyr::unnest(fit) %>%
      dplyr::mutate(sig = ifelse(p.value < 0.05, "yes", "no")) %>%
      dplyr::filter(term == "YEAR") %>%
      dplyr::select(name, sig)

    dat <- dplyr::full_join(dat,
                            pdat,
                            by = "name") %>%
      dplyr::mutate(sig = ifelse(is.na(sig), "no", sig))

    plt <- plt +
      ggplot2::geom_smooth(data = dat %>%
                             dplyr::filter(sig == "yes"),
                           method = "lm",
                           se = FALSE,
                           color = "darkgreen",
                           cex = 1.5)
  }

    if (is.null(min_year)) {
      plt <- plt +
        ggplot2::xlim(c(min(dat$YEAR), max(dat$YEAR) + 0.5))
    } else {
      plt <- plt +
        ggplot2::xlim(c(min_year, max(dat$YEAR) + 0.5))
    }
  
  # label facets ----
  finish_fig <- function() {
    if (label) {
      plt <- plt %>%
        AKesp::label_facets(open = "", close = "")
    }

    if (out == "save") {
      ggplot2::ggsave(plt, filename = paste0(name, "_page", i, ".png"), ...)
    } else if (out == "ggplot") {
      print(plt)
      cat("\n\n")
      if (silent == FALSE) {
        cat("##### Figure \\@ref(fig:traffic).", caption, "{-}")
      }
      cat("\n\n")
    } else if (out == "one_pager") {
      return(plt)
    }
    else {
      stop("Please specify output format")
    }
  }

  if (paginate == TRUE) {
    plt2 <- plt +
      ggforce::facet_wrap_paginate(~name,
                                   ncol = ncolumn,
                                   nrow = 5,
                                   scales = "free_y"
      )

    n <- ggforce::n_pages(plt2)

    for (i in 1:n) {
      plt <- plt +
        ggforce::facet_wrap_paginate(~name,
                                     ncol = ncolumn,
                                     nrow = 5,
                                     scales = "free_y",
                                     page = i
        )

      finish_fig()
    }
  } else {
    plt <- plt +
      ggplot2::facet_wrap(~name,
                          ncol = ncolumn,
                          scales = "free_y"
      )

    finish_fig()
  }
}
```

# Species-wide indicators

## Gulf Stream Anomaly
```{r, fig.height = 10}
dat <- ecodata::gsi %>%
  dplyr::mutate(Year = Time %>%
                  stringr::str_trunc(4, "right", "") %>%
                  as.numeric(),
                Month = Time %>%
                  stringr::str_trunc(2, "left", "") %>%
                  as.numeric(),
                Month = ifelse(Month == 0.1, 10, Month),
                Month = month.name[Month],
                INDICATOR_NAME = paste(Month, "GSI")) %>%
  dplyr::rename(YEAR = Time,
                DATA_VALUE = Value)

esp_traffic_fig(dat, 
                min_year = 1954,
                ncolumn = 2)
```

```{r, child = here::here("regime_child.Rmd")}
```

### Annual anomaly
```{r, fig.height = 3}
dat <- ecodata::gsi %>%
  dplyr::mutate(Year = Time %>%
                  stringr::str_trunc(4, "right", "") %>%
                  as.numeric()) %>%
  dplyr::group_by(Year) %>%
  dplyr::summarise(DATA_VALUE = mean(Value),
                   DATA_SD = sd(Value),
                   INDICATOR_NAME = "Annual GSI") %>%
  dplyr::rename(YEAR = Year)

esp_traffic_fig(dat, 
                min_year = 1954,
                sd = TRUE)
```

```{r, child = here::here("regime_child.Rmd")}
```

# Western Gulf of Maine

## Monthly Sea Surface Temperature

January and July are the most important spawning months

```{r}
dat <- read.csv(url("https://raw.githubusercontent.com/kimberly-bastille/ESPplayground/main/data-raw/cod_temperature.csv")) %>%
  dplyr::select(-X) %>%
  dplyr::distinct() %>%
  dplyr::filter(Region == "WGOM",
                Month == 1 | Month == 7) %>%
  dplyr::mutate(Month = month.name[Month]) %>%
  dplyr::rename(YEAR = Year) %>%
  dplyr::mutate(INDICATOR_NAME = paste(Month, INDICATOR_NAME))

esp_traffic_fig(dat, 
                min_year = 1982,
                ncolumn = 1,
                sd = TRUE)
```

### Use in assessment
Direct covariate on recruitment (omit July area under 11C)

## Marine heatwaves
```{r, fig.height = 3.5}
dat <- ecodata::ESP_heatwave_cod %>%
  dplyr::filter(stock_id == "WGOM") %>%
  dplyr::group_by(Time, Var) %>%
  dplyr::summarise(max_value = max(Value)) %>%
  dplyr::rename(YEAR = Time,
                INDICATOR_NAME = Var,
                DATA_VALUE = max_value)

esp_traffic_fig(dat, 
                min_year = 1982,
                ncolumn = 1)
```
```{r, child = here::here("regime_child.Rmd")}
```

### Use in assessment
Regime shift natural mortality in ~2010

## Zooplankton

Winter and summer are probably the most important seasons for larvae. 

```{r}
winter <- read.csv(here::here("WGOM_winter_zooplankton.csv"))
spring <- read.csv(here::here("WGOM_spring_zooplankton.csv"))
summer <- read.csv(here::here("WGOM_summer_zooplankton.csv"))
fall <- read.csv(here::here("WGOM_fall_zooplankton.csv"))

dat <- rbind(winter, spring, summer, fall) %>%
  dplyr::select(year, season, ctyp_100m3, calfin_100m3, pseudo_100m3, tlong_100m3) %>%
  dplyr::rename(YEAR = year) %>%
  tidyr::pivot_longer(cols = c("ctyp_100m3", "calfin_100m3", "pseudo_100m3", "tlong_100m3")) %>%
  dplyr::mutate(INDICATOR_NAME = paste(season, name)) %>%
  dplyr::rename(DATA_VALUE = value)

dat <- dat %>%
  dplyr::filter(stringr::str_detect(INDICATOR_NAME, "Winter") |
                  stringr::str_detect(INDICATOR_NAME, "Summer"))
```

### Calanus
```{r, fig.height = 3.5}
esp_traffic_fig(dat %>%
                  dplyr::filter(name == "calfin_100m3"), 
                min_year = 1977,
                ncolumn = 1)
```

### Pseudocalanus
```{r, fig.height = 3.5}
esp_traffic_fig(dat %>%
                  dplyr::filter(name == "pseudo_100m3"), 
                min_year = 1977,
                ncolumn = 1)
```

### Centropages
```{r, fig.height = 3.5}
esp_traffic_fig(dat %>%
                  dplyr::filter(name == "ctyp_100m3"), 
                min_year = 1977,
                ncolumn = 1)
```

### Temora
```{r, fig.height = 3.5}
esp_traffic_fig(dat %>%
                  dplyr::filter(name == "tlong_100m3"), 
                min_year = 1977,
                ncolumn = 1)
```

### Use in assessment
Direct covariate on recruitment?

## Condition
```{r}
dat <- read.csv(here::here("Condition/ADIOS_SV_164712_WGOM_NONE_relative_k.csv")) %>%
  dplyr::mutate(INDICATOR_NAME = paste(SEASON, SURVEY)) %>%
  dplyr::select(YEAR, INDICATOR_NAME, K_rel) %>%
  dplyr::group_by(YEAR, INDICATOR_NAME) %>%
  dplyr::summarise(DATA_VALUE = mean(K_rel),
                   DATA_SD = sd(K_rel))

esp_traffic_fig(dat, 
                min_year = 1992,
                ncolumn = 1,
                sd = TRUE)
```

```{r, child = here::here("regime_child.Rmd")}
```

### Use in assessment
?

# Georges Bank

## Monthly Sea Surface Temperature

April is the most important spawning month

```{r, fig.height = 3.5}
dat <- read.csv(url("https://raw.githubusercontent.com/kimberly-bastille/ESPplayground/main/data-raw/cod_temperature.csv")) %>%
  dplyr::filter(Region == "GBK",
                Month == 4) %>%
  dplyr::mutate(Month = month.name[Month]) %>%
  dplyr::rename(YEAR = Year) %>%
  dplyr::mutate(INDICATOR_NAME = paste(Month, INDICATOR_NAME))

esp_traffic_fig(dat, 
                min_year = 1982,
                ncolumn = 1,
                sd = TRUE)
```

### Use in assessment
Direct covariate on recruitment

## Marine heatwaves
```{r, fig.height = 3.5}
dat <- ecodata::ESP_heatwave_cod %>%
  dplyr::filter(stock_id == "GBK") %>%
  dplyr::group_by(Time, Var) %>%
  dplyr::summarise(max_value = max(Value)) %>%
  dplyr::rename(YEAR = Time,
                INDICATOR_NAME = Var,
                DATA_VALUE = max_value)

esp_traffic_fig(dat, 
                min_year = 1982,
                ncolumn = 1)
```

```{r, child = here::here("regime_child.Rmd")}
```

### Use in assessment
Regime shift natural mortality in ~2010

## Zooplankton

Spring is probably the most important season for larvae.

```{r}
winter <- read.csv(here::here("GB_winter_zooplankton.csv"))
spring <- read.csv(here::here("GB_spring_zooplankton.csv"))
summer <- read.csv(here::here("GB_summer_zooplankton.csv"))
fall <- read.csv(here::here("GB_fall_zooplankton.csv"))

dat <- rbind(winter, spring, summer, fall) %>%
  dplyr::select(year, season, ctyp_100m3, calfin_100m3, pseudo_100m3, tlong_100m3) %>%
  dplyr::rename(YEAR = year) %>%
  tidyr::pivot_longer(cols = c("ctyp_100m3", "calfin_100m3", "pseudo_100m3", "tlong_100m3")) %>%
  dplyr::mutate(INDICATOR_NAME = paste(season, name)) %>%
  dplyr::rename(DATA_VALUE = value)

dat <- dat %>%
  dplyr::filter(stringr::str_detect(INDICATOR_NAME, "Spring"))
```

### Calanus
```{r, fig.height = 2}
esp_traffic_fig(dat %>%
                  dplyr::filter(name == "calfin_100m3"), 
                min_year = 1977,
                ncolumn = 1)
```

### Pseudocalanus
```{r, fig.height = 2}
esp_traffic_fig(dat %>%
                  dplyr::filter(name == "pseudo_100m3"), 
                min_year = 1977,
                ncolumn = 1)
```

### Centropages
```{r, fig.height = 2}
esp_traffic_fig(dat %>%
                  dplyr::filter(name == "ctyp_100m3"), 
                min_year = 1977,
                ncolumn = 1)
```

### Temora
```{r, fig.height = 2}
esp_traffic_fig(dat %>%
                  dplyr::filter(name == "tlong_100m3"), 
                min_year = 1977,
                ncolumn = 1)
```

### Use in assessment


## Condition
```{r, fig.height = 3.5}
dat <- read.csv(here::here("Condition/ADIOS_SV_164712_GBK_NONE_relative_k.csv")) %>%
  dplyr::mutate(INDICATOR_NAME = paste(SEASON, SURVEY)) %>%
  dplyr::select(YEAR, INDICATOR_NAME, K_rel) %>%
  dplyr::group_by(YEAR, INDICATOR_NAME) %>%
  dplyr::summarise(DATA_VALUE = mean(K_rel),
                   DATA_SD = sd(K_rel))

esp_traffic_fig(dat, 
                min_year = 1992,
                ncolumn = 1,
                sd = TRUE)
```

```{r, child = here::here("regime_child.Rmd")}
```

### Use in assessment
?

# Southern New England

## Monthly Sea Surface Temperature

April is the most important spawning month

```{r, fig.height = 3.5}
dat <- read.csv(url("https://raw.githubusercontent.com/kimberly-bastille/ESPplayground/main/data-raw/cod_temperature.csv")) %>%
  dplyr::filter(Region == "SNE",
                Month == 4) %>%
  dplyr::mutate(Month = month.name[Month]) %>%
  dplyr::rename(YEAR = Year) %>%
  dplyr::mutate(INDICATOR_NAME = paste(Month, INDICATOR_NAME))

esp_traffic_fig(dat, 
                min_year = 1982,
                ncolumn = 1,
                sd = TRUE)
```

### Use in assessment
?

## Marine heatwaves
```{r, fig.height = 3.5}
dat <- ecodata::ESP_heatwave_cod %>%
  dplyr::filter(stock_id == "SNE") %>%
  dplyr::group_by(Time, Var) %>%
  dplyr::summarise(max_value = max(Value)) %>%
  dplyr::rename(YEAR = Time,
                INDICATOR_NAME = Var,
                DATA_VALUE = max_value)

esp_traffic_fig(dat, 
                min_year = 1982,
                ncolumn = 1)
```

```{r, child = here::here("regime_child.Rmd")}
```

### Use in assessment
Regime shift natural mortality in ~2010

## Zooplankton

Spring is probably the most important season for larvae.

```{r}
winter <- read.csv(here::here("SNE_winter_zooplankton.csv"))
spring <- read.csv(here::here("SNE_spring_zooplankton.csv"))
summer <- read.csv(here::here("SNE_summer_zooplankton.csv"))
fall <- read.csv(here::here("SNE_fall_zooplankton.csv"))

dat <- rbind(winter, spring, summer, fall) %>%
  dplyr::select(year, season, ctyp_100m3, calfin_100m3, pseudo_100m3, tlong_100m3) %>%
  dplyr::rename(YEAR = year) %>%
  tidyr::pivot_longer(cols = c("ctyp_100m3", "calfin_100m3", "pseudo_100m3", "tlong_100m3")) %>%
  dplyr::mutate(INDICATOR_NAME = paste(season, name)) %>%
  dplyr::rename(DATA_VALUE = value)

dat <- dat %>%
  dplyr::filter(stringr::str_detect(INDICATOR_NAME, "Spring"))
```

### Calanus
```{r, fig.height = 2}
esp_traffic_fig(dat %>%
                  dplyr::filter(name == "calfin_100m3"), 
                min_year = 1977,
                ncolumn = 1)
```

### Pseudocalanus
```{r, fig.height = 2}
esp_traffic_fig(dat %>%
                  dplyr::filter(name == "pseudo_100m3"), 
                min_year = 1977,
                ncolumn = 1)
```

### Centropages
```{r, fig.height = 2}
esp_traffic_fig(dat %>%
                  dplyr::filter(name == "ctyp_100m3"), 
                min_year = 1977,
                ncolumn = 1)
```

### Temora
```{r, fig.height = 2}
esp_traffic_fig(dat %>%
                  dplyr::filter(name == "tlong_100m3"), 
                min_year = 1977,
                ncolumn = 1)
```

### Use in assessment


## Condition
```{r, fig.height = 5}
dat <- read.csv(here::here("Condition/ADIOS_SV_164712_SNEMA_NONE_relative_k.csv")) %>%
  dplyr::mutate(INDICATOR_NAME = paste(SEASON, SURVEY)) %>%
  dplyr::select(YEAR, INDICATOR_NAME, K_rel) %>%
  dplyr::group_by(YEAR, INDICATOR_NAME) %>%
  dplyr::summarise(DATA_VALUE = mean(K_rel),
                   DATA_SD = sd(K_rel))

esp_traffic_fig(dat, 
                min_year = 1992,
                ncolumn = 1,
                sd = TRUE)
```

```{r, child = here::here("regime_child.Rmd")}
```

### Use in assessment
?

# Eastern Gulf of Maine

## Monthly Sea Surface Temperature

January and July are the most important spawning months

```{r}
dat <- read.csv(url("https://raw.githubusercontent.com/kimberly-bastille/ESPplayground/main/data-raw/cod_temperature.csv")) %>%
  dplyr::filter(Region == "EGOM",
                Month == 1 | Month == 7) %>%
  dplyr::mutate(Month = month.name[Month]) %>%
  dplyr::rename(YEAR = Year) %>%
  dplyr::mutate(INDICATOR_NAME = paste(Month, INDICATOR_NAME))

esp_traffic_fig(dat, 
                min_year = 1982,
                ncolumn = 1,
                sd = TRUE)
```

## Marine heatwaves
```{r, fig.height = 3.5}
dat <- ecodata::ESP_heatwave_cod %>%
  dplyr::filter(stock_id == "EGOM") %>%
  dplyr::group_by(Time, Var) %>%
  dplyr::summarise(max_value = max(Value)) %>%
  dplyr::rename(YEAR = Time,
                INDICATOR_NAME = Var,
                DATA_VALUE = max_value)

esp_traffic_fig(dat, 
                min_year = 1982,
                ncolumn = 1)
```

```{r, child = here::here("regime_child.Rmd")}
```

### Use in assessment
Regime shift natural mortality in ~2010

## Zooplankton

Winter and summer are probably the most important seasons for larave. 

```{r}
winter <- read.csv(here::here("EGOM_winter_zooplankton.csv"))
spring <- read.csv(here::here("EGOM_spring_zooplankton.csv"))
summer <- read.csv(here::here("EGOM_summer_zooplankton.csv"))
fall <- read.csv(here::here("EGOM_fall_zooplankton.csv"))

dat <- dplyr::full_join(winter, spring) %>%
  dplyr::full_join(summer) %>%
  dplyr::full_join(fall) %>%
  dplyr::select(year, season, ctyp_100m3, calfin_100m3, pseudo_100m3, tlong_100m3) %>%
  dplyr::rename(YEAR = year) %>%
  tidyr::pivot_longer(cols = c("ctyp_100m3", "calfin_100m3", "pseudo_100m3", "tlong_100m3")) %>%
  dplyr::mutate(INDICATOR_NAME = paste(season, name)) %>%
  dplyr::rename(DATA_VALUE = value)

dat <- dat %>%
  dplyr::filter(stringr::str_detect(INDICATOR_NAME, "Winter") |
                  stringr::str_detect(INDICATOR_NAME, "Summer"))
```

### Calanus
```{r, fig.height = 3.5}
esp_traffic_fig(dat %>%
                  dplyr::filter(name == "calfin_100m3"), 
                min_year = 1977,
                ncolumn = 1)
```

### Pseudocalanus
```{r, fig.height = 3.5}
esp_traffic_fig(dat %>%
                  dplyr::filter(name == "pseudo_100m3"), 
                min_year = 1977,
                ncolumn = 1)
```

### Centropages
```{r, fig.height = 3.5}
esp_traffic_fig(dat %>%
                  dplyr::filter(name == "ctyp_100m3"), 
                min_year = 1977,
                ncolumn = 1)
```

### Temora
```{r, fig.height = 3.5}
esp_traffic_fig(dat %>%
                  dplyr::filter(name == "tlong_100m3"), 
                min_year = 1977,
                ncolumn = 1)
```

```{r, child = here::here("regime_child.Rmd")}
```

## Condition
```{r, fig.height = 3.5}
dat <- read.csv(here::here("Condition/ADIOS_SV_164712_EGOM_NONE_relative_k.csv")) %>%
  dplyr::mutate(INDICATOR_NAME = paste(SEASON, SURVEY)) %>%
  dplyr::select(YEAR, INDICATOR_NAME, K_rel) %>%
  dplyr::group_by(YEAR, INDICATOR_NAME) %>%
  dplyr::summarise(DATA_VALUE = mean(K_rel),
                   DATA_SD = sd(K_rel))

esp_traffic_fig(dat, 
                min_year = 1992,
                ncolumn = 1,
                sd = TRUE)
```

```{r, child = here::here("regime_child.Rmd")}
```

